---
title: "Code_EDA"
author: "Madhura Jadhav"
output: html_document
---


```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)

library(skimr)
library(summarytools)
library(lubridate)
library(stringr)
library(Hmisc)
library(broom)
library(car)

library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(xgboost)
library(vtreat)
```


```{r}
original_data = read.csv('analysisData.csv')
dim(original_data)
```

### Filtering the data to remove out of scope / redundant attributes
```{r}
data=original_data[c('id','host_name', 'host_since', 'host_response_time','host_response_rate', 'host_is_superhost', 'host_total_listings_count', 'neighbourhood_cleansed', 'neighbourhood_group_cleansed', 'zipcode', 'property_type', 'room_type', 'accommodates', 'bathrooms', 'bedrooms', 'beds', 'amenities', 'price','cleaning_fee','guests_included','extra_people', 'minimum_nights','maximum_nights','availability_30','availability_60','availability_90','availability_365', 'number_of_reviews','number_of_reviews_ltm','first_review','last_review','review_scores_rating', 'review_scores_accuracy', 'review_scores_cleanliness', 'review_scores_checkin', 'review_scores_communication', 'review_scores_location', 'review_scores_value', 'instant_bookable','calculated_host_listings_count','reviews_per_month')]
dim(data)

```
```{r}
#library(summarytools)
print(dfSummary(data,style='grid',graph.col = T),method = 'render')
```


## Missing Values

#### Missing Values Columns: cleaning_fee 5276,  beds 44, host_total_listings_count 21, reviews_per_month 2, host_response_time 8679, host_response_rate8679
#### Will drop rows with missing values <50 for a column 

```{r}
data = data %>% drop_na(reviews_per_month)
data = data %>% drop_na(host_total_listings_count)
data = data %>% drop_na(beds)
```

#### Cleaning fees is option so absence of it is cleaning fee=0
```{r}
data = data %>% mutate(cleaning_fee = ifelse(is.na(data$cleaning_fee),0,data$cleaning_fee))
```

#### We will drop host_response_time 8679, host_response_rate 8679 as we have ~25% missing data
```{r}
data = data[ , -which(names(data) %in% c("host_response_rate","host_response_time"))]
```

## Data Transformations

### Converting columns with logical data type to 1 and 0 

#### host_is_superhost,instant_bookable 

```{r}
data$host_is_superhost = as.numeric(ifelse(data$host_is_superhost == 't', 1, 0))
data$instant_bookable = as.numeric(ifelse(data$instant_bookable == 't', 1, 0))
```

### New Features

#### host_since can be used to calculate host active duration in months
#### first_review & last_review can be used to calculate listing active duration in months
#### we can drop column last_review,first_review, host_since after calculating new features

```{r}
#max(data$last_review) is '2020-01-03'

data$listing_active_duration = round(as.numeric(difftime(data$last_review, data$first_review, units = "days"))/(365.25/12),2)
data$host_active_duration = round(as.numeric(difftime('2020-01-03', data$host_since, units ="days"))/(365.25/12),2)

data = subset(data, select = -c(host_since,first_review,last_review) )
```

#### Amenities can be extracted to check important individual amenity
```{r}
#library(stringr)
data$TV = str_extract(data$amenities, "TV")
data$TV[is.na(data$TV)] <- 0
data$TV[data$TV == "TV"] <- 1

data$Elevator = str_extract(data$amenities, "Elevator")
data$Elevator[is.na(data$Elevator)] <- 0
data$Elevator[data$Elevator == "Elevator"] <- 1
```

### Combine variable levels

#### property_type
##### combine levels with <1% data to 'Others' making total number of levels for property_type as 6

```{r}
data %>% group_by(property_type) %>% summarise(n=n()) %>% arrange(desc(n))
```
```{r}
data=data %>% mutate(property_type = fct_lump(property_type, prop = 0.01)) 
```

#### room_type
##### updating 'Hotel room' level as 'Private room' as we have just 3 'Hotel room' records 

```{r}
data$room_type = fct_collapse(data$room_type, 'Private room' = c('Hotel room', 'Private room'))
```

#### bathrooms
##### updating 0.5 bathroom levels and reducing number of levels
```{r}
data$bathrooms[data$bathrooms < 1] <- 1  
data$bathrooms[ data$bathrooms == 1.5] <- 2  
data$bathrooms[ data$bathrooms == 2.5] <- 3  
data$bathrooms[ data$bathrooms == 3.5] <- 4  
data$bathrooms[ data$bathrooms > 4] <- 4 
```

#### bedrooms
##### reducing number of levels from 12 to 7
```{r}
data$bedrooms[data$bedrooms > 6] <- 6  
```

## Delete erroneous data

#### price
##### delete listings with price=0,assuming it is data error
```{r}
data= data[!(data$price==0),]
```

#### zipcode
##### clean dirty zipcode data, treat NAs with mode value
```{r}
data$zipcode=str_replace_all(data$zipcode,"NY ","")
data$zipcode=str_replace_all(data$zipcode," ","")
data$zipcode[data$zipcode=="11385-2308"] = '11385'
data$zipcode[data$zipcode=="11103-3233"] = '11103'
data$zipcode[data$zipcode=="10003-8623"] = '10003'
data$zipcode[data$zipcode=="11413-3220"] = '11413'
data$zipcode[data$zipcode=="10065"] = '10021'
data$zipcode[data$zipcode=="11249"] = '11211'
data$zipcode <- gsub('NA','11211',data$zipcode)
```

```{r}
# odata %>% 
#    filter(neighbourhood_group_cleansed=='Staten Island') %>%
#    count(zipcode, sort = TRUE) 

#Brooklyn -11211
#Bronx - 10469
#Manhattan - 10002
#Queens- 11385
#SA - 10301
```

# EDA


##Target Variable - price distribution
```{r}
ggplot(data, aes(x=price)) + 
  geom_histogram(binwidth = 10)
```
#### will be analyzing numeric variables across price with scatter plots
#### will be analyzing categorical and few numeric variables across price with box plots

```{r}
box = geom_boxplot(varwidth=T) 
scatter = geom_point()
```

# Analyzing Categorical Variable
## categorical: neighbourhood_cleansed,neighbourhood_group_cleansed

### neighbourhood_cleansed -- explored neighbourhood_cleansed for each borough.. but Too many levels and no so DROP

```{r}
# data %>% 
#   filter(neighbourhood_group_cleansed == 'Brooklyn') %>% 
#   ggplot(aes(x=neighbourhood_cleansed,y=price)) + box
```
```{r}
# data %>% 
#   filter(neighbourhood_group_cleansed == 'Bronx') %>% 
#   ggplot(aes(x=neighbourhood_cleansed,y=price)) + box
```
```{r}
# data %>% 
#   filter(neighbourhood_group_cleansed == 'Manhattan') %>% 
#   ggplot(aes(x=neighbourhood_cleansed,y=price)) + box
```
```{r}
# data %>% 
#   filter(neighbourhood_group_cleansed == 'Queens') %>% 
#   ggplot(aes(x=neighbourhood_cleansed,y=price)) + box
```
```{r}
# data %>% 
#   filter(neighbourhood_group_cleansed == 'Staten Island') %>% 
#   ggplot(aes(x=neighbourhood_cleansed,y=price)) + box
```


### neighbourhood_group_cleansed

```{r}
ggplot(data, aes(x=neighbourhood_group_cleansed,y=price)) + box
```

# Analyzing Factor Variables
## factor: property_type, room_type

### property_type

```{r}
ggplot(data, aes(x=property_type,y=price)) + box
```

### room_type

```{r}
ggplot(data, aes(x=room_type,y=price)) + box
```

# Analyzing Numeric Factor Variables
## numeric factor: host_is_superhost, accommodates, bathrooms, bedrooms, beds, availability_30, review_scores_rating, review_scores_accuracy,  review_scores_cleanliness, review_scores_checkin, review_scores_communication,  review_scores_location,  review_scores_value, instant_bookable

### host_is_superhost
```{r}
ggplot(data, aes(x=as.factor(host_is_superhost),y=price)) + box
```

### accommodates -- might convert 9 onwards as 9
```{r}
ggplot(data, aes(x=as.factor(accommodates),y=price)) + box
```
### bathrooms- we have converted 

```{r}
ggplot(data, aes(x=as.factor(bathrooms),y=price)) + box
```

### bedrooms

```{r}
ggplot(data, aes(x=as.factor(bedrooms),y=price)) + box
```

###beds -- might convert 8 onwards as 8

```{r}
ggplot(data, aes(x=as.factor(beds),y=price)) + box
```
### availability_30

```{r}
ggplot(data, aes(x=as.factor(availability_30),y=price)) + box
```
### review_scores_rating

```{r}
data%>%
  mutate(review_scores_rating_bins = cut(review_scores_rating, breaks = c(0,10,20,30,40,50,60,70,80,90,100))) %>% 
  ggplot(aes(review_scores_rating_bins,price)) + box

```



### review_scores_accuracy

```{r}
ggplot(data, aes(x=as.factor(review_scores_accuracy),y=price)) + box
```

### review_scores_cleanliness

```{r}
ggplot(data, aes(x=as.factor(review_scores_cleanliness),y=price)) + box
```

### review_scores_checkin

```{r}
ggplot(data, aes(x=as.factor(review_scores_checkin),y=price)) + box
```

### review_scores_communication

```{r}
ggplot(data, aes(x=as.factor(review_scores_communication),y=price)) + box
```

### review_scores_location

```{r}
ggplot(data, aes(x=as.factor(review_scores_location),y=price)) + box
```

### review_scores_value

```{r}
ggplot(data, aes(x=as.factor(review_scores_value),y=price)) + box
```

### instant_bookable

```{r}
ggplot(data, aes(x=as.factor(instant_bookable),y=price)) + box
```
# Analyzing Numeric Variables
## numeric: host_total_listings_count, cleaning_fee, minimum_nights, maximum_nights, number_of_reviews, number_of_reviews_ltm, calculated_host_listings_count, reviews_per_month, listing_active_duration, host_active_duration

### host_total_listings_count

```{r}
ggplot(data, aes(x=host_total_listings_count,y=price)) + scatter
```

### cleaning_fee

```{r}
ggplot(data, aes(x=log(cleaning_fee),y=price)) + scatter #+ geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1))
```

### minimum_nights

```{r}
ggplot(data, aes(x=(minimum_nights),y=price)) + scatter
```

### maximum_nights

```{r}
ggplot(data, aes(x=maximum_nights,y=price)) + scatter
```
### number_of_reviews

```{r}
ggplot(data, aes(x=number_of_reviews,y=price)) + scatter
```

### number_of_reviews_ltm

```{r}
ggplot(data, aes(x=number_of_reviews_ltm,y=price)) + scatter
```


### calculated_host_listings_count

```{r}
ggplot(data, aes(x=calculated_host_listings_count,y=price)) + scatter
```

### reviews_per_month

```{r}
ggplot(data, aes(x=reviews_per_month,y=price)) + scatter
```

### listing_active_duration

```{r}
ggplot(data, aes(x=listing_active_duration,y=price)) + scatter
```

### host_active_duration

```{r}
ggplot(data, aes(x=host_active_duration,y=price)) + scatter
```

## Correlations
```{r}
#library(corrplot)
#numeric variables
numericVars <- which(sapply(data, is.numeric)) 
data_numVar <- data[, numericVars]

#correlations of all numeric variables
cor_numVar <- cor(data_numVar, use="pairwise.complete.obs")

#sorting on decreasing correlations with price
cor_sorted <- as.matrix(sort(cor_numVar[,'price'], decreasing = TRUE))
print(cor_sorted)
```

#Multivariate Filter
```{r}
model = lm(price~.,data_numVar)
#library(broom)
summary(model) %>%
  tidy()
```
```{r}
#library(car)
vif(model)
```
```{r}
data.frame(Predictor = names(vif(model)), VIF = vif(model)) %>%
  ggplot(aes(x=VIF, y = reorder(Predictor, VIF), fill=VIF))+
  geom_col()+
  geom_vline(xintercept=5, color = 'gray', size = 1.5)+
  geom_vline(xintercept = 10, color = 'red', size = 1.5)+
  scale_fill_gradient(low = '#fff7bc', high = '#d95f0e')+
  scale_y_discrete(name = "Predictor")+
  scale_x_continuous(breaks = seq(5,30,5))+
  theme_classic()
```

##Write transformed data to csv file

```{r}
write.csv(data, "clean_full_analysis_data.csv",row.names = F)
```


##Write selective features transformed data to csv file

```{r}
data_trimmed = data[,c('price','zipcode','room_type','bedrooms','accommodates','neighbourhood_group_cleansed','availability_30','host_is_superhost','review_scores_rating','review_scores_location','TV','Elevator','cleaning_fee','property_type','minimum_nights','bathrooms')]

write.csv(data_trimmed, "clean_trimmed_analysis_data.csv",row.names = F)
```
